\bf Landslide: \\ Systematic Dynamic Race Detection in Kernel Space

### 2012-05-17 evaluation.tex
eval numbers bw print support
change figs to tables
add eval graphs

\newcommand\asdf[1]{\hspace{-0.05in}\footnotesize\bf{#1}\hspace{-0.05in}}
\revision{
\begin{figure}[t]
	\begin{center}
	\small
	\begin{tabular}{|l||c|c|c|c|c|c||c|c|}
		\hline
		\multirow{2}{*}{\bf Group} & \multicolumn{6}{|c||}{\bf Minutes spent doing\dots} & \multicolumn{2}{|c|}{\bf Total time spent} \\
		\cline{2-9}
		& \asdf{annotating} & \asdf{config} & \asdf{student.c} & \asdf{fixing} & \asdf{customising} & \asdf{found races} & \asdf{required} & \asdf{refinement} \\
		\hline \hline
		% ntaha/jmchow
		\multirow{2}{*}{Group 1} & 35 & {\footnotesize \em N/A} & {\footnotesize \em N/A} & 5 & {\footnotesize \em N/A} & {\footnotesize \em N/A} & {\footnotesize \em N/A} & {\footnotesize \em N/A} \\
		\cline{2-9}
		& 5 & 30 & 10 & 100 & {\footnotesize \em N/A} & {\footnotesize \em N/A} & 145 & {\footnotesize \em N/A} \\
		\hline
		% echo/margaret
		\multirow{2}{*}{Group 2} & 20 & 30 & 10 & 80 & {\footnotesize \em N/A} & 10 & 140 & 10 \\
		\cline{2-9}
		& 15 & 45 & 30 & 60 & {\footnotesize \em N/A} & 30 & 150 & 30 \\
		\hline
		% tpassaro
		\multirow{2}{*}{Group 3} & 15 & 10 & 5 & 60 & 30 & 30 & 90 & 60 \\
		\cline{2-9}
		& 15 & 10 & 5 & 60 & 30 & 30 & 90 & 60 \\
		\hline
		% pjumde
		\multirow{2}{*}{Group 4} & 10 & 42 & 3 & 5 & 30 & {\footnotesize \em N/A} & 60 & 30 \\
		\cline{2-9}
		& 15 & 35 & {\footnotesize \em N/A} & 5 & 35 & {\footnotesize \em N/A} & {\footnotesize \em N/A} & 35 \\
		\hline
		% apodolsk
		Group 5 & 50 & 15 & {\footnotesize \em N/A} & 15 & {\footnotesize \em N/A} & {\footnotesize \em N/A} & {\footnotesize \em N/A} & {\footnotesize \em N/A} \\
		\hline
		Former TA & 40 & 20 & 3 & 95 & {\footnotesize \em N/A} & 30 & 158 & 30 \\
		\hline
		\hline
		{\bf Average } & {\bf 22 } & {\bf 26.33 } & {\bf 9.43 } & {\bf 48.5 } & {\bf 31.25 } & {\bf 26 } & {\bf 119 } & {\bf 36.43} \\
		\hline
	\end{tabular}
	\end{center}
	\caption{Self-reported times spent by users on each phase while using Landslide. ``N/A'' indicates a student reported spending no time on a particular phase. ``Required'' indicates the sum of the times from the first four phases; ``refinement'' inducates the sum from the last two phases.
	``N/A'' indicates that a user did not report spending any time on a phase; the ``required'' field is ``N/A'' if any of its represented phases are, and ``N/A'' fields are not reflected in the averages.}
	\label{fig:student-times}
\end{figure}

@@

Table~\ref{fig:student-times} presents the time breakdown among phases that students experienced while using Landslide. On average, students spent two hours getting Landslide to work with their kernels, and slightly more than half an hour customising the search and/or studying found bugs on top of that.

@@

Table~\ref{fig:numbers} shows the time it takes to find each of these bugs using Landslide, configured with several different sets of decision points, and using conventional stress testing.

\newcommand\bugnum[2]{\textcolor{BrickRed}{{\bf #1} {\scriptsize \em (#2)}}}
\newcommand\nobugnum[2]{\textcolor{Blue}{{\em #1} {\scriptsize \em (#2)}}}

\begin{figure*}[t!]
	\begin{center}
	\small
	\begin{tabular}{|l||c|c|c|c||c|}
		\hline
		\multirow{2}{*}{\bf Kernel and test case} & \multicolumn{4}{c||}{\bf Decision points used} & \multirow{2}{*}{\bf Stress test} \\
		\cline{2-5}
		& \bf default & \bf lock & \bf unlock & \bf both & \\
		\hline\hline
		POBBLES vanish/vanish (a) & \nobugnum{31.8}{0.6} & \bugnum{57.1}{1.7} & $\infty$ & $\infty$ & \\
		\hline
		POBBLES vanish/vanish (b) & \nobugnum{32.0}{0.4} & \bugnum{51.5}{2.1} & \nobugnum{8057.9}{336.9} & $\infty$ & \\
		\hline
		POBBLES wait/wait & \bugnum{23.3}{0.7} & \bugnum{27.9}{0.8} & \bugnum{27.9}{2.1} & \bugnum{41.6}{1.4} & \\
		\hline
		POBBLES thread\_fork/vanish & \bugnum{22.0}{0.6} & \bugnum{37.4}{1.1} & \bugnum{27.6}{0.5} & \bugnum{72.0}{2.6} & \\
		\hline
		LudicrOS vanish/vanish & \nobugnum{13.2}{0.2} & \bugnum{13.7}{0.7} & \bugnum{34.6}{1.1} & \bugnum{17.1}{0.3} & \\
		\hline
		LudicrOS yield/vanish & \nobugnum{12.3}{0.3} & \bugnum{11.4}{0.4} & \nobugnum{27.4}{0.8} & \bugnum{11.7}{0.4} & \\
		\hline
		\multicolumn{6}{c}{Key: \bugnum{seconds}{stddev} indicates bug found; \nobugnum{seconds}{stddev} indicates whole tree explored with no bug.} \\
	\end{tabular}
	\end{center}
	\caption{Comparison of time taken (in seconds) to find bugs using Landslide and using conventional stress testing.
	Landslide's times are given for several different sets of decision points: the default set, consisting only of voluntary reschedules (Section~\ref{sec:components-arbiter}); and using custom decision points in addition to the default set: calls to \texttt{mutex\_lock}, calls to \texttt{mutex\_unlock}, and both.
	All numbers represent the average from 5 trials, with the standard deviations given in parentheses. ``$\infty$'' indicates the bug was not found with stress testing (after 1 hour), or that Landslide's search did not finish (after 8 hours). % TODO: what timeout for stress testing?
	}
	\label{fig:numbers}
\end{figure*}

% TODO: have a graph

The Landslide experimental set-up is as follows:

@@

%The stress test experimental set-up is as follows:
%
%\begin{itemize}
%	\item A wrapper program runs 1024 simultaneous copies of the same test program that Landslide uses. Whenever one of them exits, the wrapper starts a new one.
%	\item All stress test trials are on on the ``412 lab crashbox'' (3.2 GHz Pentium D).
%	\item The kernel and test case are configured to not print any messages during the test
%\end{itemize}
% TODO: say specs of crash machine

@@

\newcommand\bugtree[1]{\textcolor{BrickRed}{\bf #1}}
\newcommand\nobugtree[1]{\textcolor{Blue}{\em #1}}
\begin{figure*}[t!]
	\begin{center}
	\small
	\begin{tabular}{|l|l||c|c|c|c|}
		\hline
		\multirow{2}{*}{\bf Kernel and test case} & \multirow{2}{*}{\bf Property of tree} & \multicolumn{4}{|c|}{\bf Decision points used} \\
		\cline{3-6}
		& & \bf default & \bf lock & \bf unlock & \bf both \\
		\hline\hline
		\multirow{4}{*}{POBBLES vanish/vanish (a)} & Decision points & \nobugtree{56} & \bugtree{1296} & N/A & N/A \\
		& Total backtracks   & \nobugtree{16} & \bugtree{376} & N/A & N/A \\
		& Average branch depth & \nobugtree{5} & \bugtree{19} & N/A & N/A \\
		\hline
		\multirow{4}{*}{POBBLES vanish/vanish (b)} & Decision points & \nobugtree{56} & \bugtree{1295} & \nobugtree{382071} & N/A \\
		& Total backtracks   & \nobugtree{16} & \bugtree{376} & \nobugtree{112706} & N/A \\
		& Average branch depth & \nobugtree{5} & \bugtree{17} & \nobugtree{16} & N/A \\
		\hline
		\multirow{4}{*}{POBBLES wait/wait} & Decision points & \bugtree{23} & \bugtree{102} & \bugtree{74} & \bugtree{378} \\
		& Total backtracks   & \bugtree{4} & \bugtree{17} & \bugtree{12} & \bugtree{56} \\
		& Average branch depth & \bugtree{6} & \bugtree{10} & \bugtree{9} & \bugtree{14} \\
		\hline
		\multirow{4}{*}{POBBLES thread\_fork/vanish} & Decision points & \bugtree{24} & \bugtree{394} & \bugtree{273} & \bugtree{2269} \\
		& Total backtracks   & \bugtree{5} & \bugtree{70} & \bugtree{56} & \bugtree{410} \\
		& Average branch depth & \bugtree{5} & \bugtree{16} & \bugtree{14} & \bugtree{23} \\
		\hline
		\multirow{4}{*}{LudicrOS vanish/vanish} & Decision points & \nobugtree{10} & \bugtree{16} & \bugtree{141} & \bugtree{42} \\
		& Total backtracks   & \nobugtree{2} & \bugtree{3} & \bugtree{48} & \bugtree{10} \\
		& Average branch depth & \nobugtree{2} & \bugtree{7} & \bugtree{9} & \bugtree{14} \\
		\hline
		\multirow{4}{*}{LudicrOS yield/vanish} & Decision points & \nobugtree{8} & \bugtree{5} & \nobugtree{149} & \bugtree{7} \\
		& Total backtracks   & \nobugtree{1} & \bugtree{0} & \nobugtree{43} & \bugtree{0} \\
		& Average branch depth & \nobugtree{2} & \bugtree{0} & \nobugtree{9} & \bugtree{0} \\
		\hline
	\end{tabular}
	\end{center}
	\caption{Information about the decision trees explored when finding bugs. As in the previous table, each test case was run with the four different sets of decision points. ``\nobugtree{X}'' means the tree was completely explored because Landslide did not find a bug in that configuration. ``\bugtree{X}'' reflects the portion of the tree that was explored before a bug was found.}
	\label{fig:trees}
\end{figure*}

>>>>

\newcommand\asdf[1]{\hspace{-0.05in}\footnotesize\bf{#1}\hspace{-0.05in}}
\revision{
\begin{table}[t]
	\begin{center}
	\small
	\begin{tabular}{|l||c|c|c|c|c|c||c|c|}
		\hline
		\multirow{2}{*}{\bf Group} & \multicolumn{6}{|c||}{\bf Minutes spent doing\dots} & \multicolumn{2}{|c|}{\bf Total time spent} \\
		\cline{2-9}
		& \asdf{annotating} & \asdf{config} & \asdf{student.c} & \asdf{fixing} & \asdf{customising} & \asdf{found races} & \asdf{required} & \asdf{refinement} \\
		\hline \hline
		% ntaha/jmchow
		\multirow{2}{*}{Group 1} & 35 & {\footnotesize \em N/A} & {\footnotesize \em N/A} & 5 & {\footnotesize \em N/A} & {\footnotesize \em N/A} & {\footnotesize \em N/A} & {\footnotesize \em N/A} \\
		\cline{2-9}
		& 5 & 30 & 10 & 100 & {\footnotesize \em N/A} & {\footnotesize \em N/A} & 145 & {\footnotesize \em N/A} \\
		\hline
		% echo/margaret
		\multirow{2}{*}{Group 2} & 20 & 30 & 10 & 80 & {\footnotesize \em N/A} & 10 & 140 & 10 \\
		\cline{2-9}
		& 15 & 45 & 30 & 60 & {\footnotesize \em N/A} & 30 & 150 & 30 \\
		\hline
		% tpassaro
		\multirow{2}{*}{Group 3} & 15 & 10 & 5 & 60 & 30 & 30 & 90 & 60 \\
		\cline{2-9}
		& 15 & 10 & 5 & 60 & 30 & 30 & 90 & 60 \\
		\hline
		% pjumde
		\multirow{2}{*}{Group 4} & 10 & 42 & 3 & 5 & 30 & {\footnotesize \em N/A} & 60 & 30 \\
		\cline{2-9}
		& 15 & 35 & {\footnotesize \em N/A} & 5 & 35 & {\footnotesize \em N/A} & {\footnotesize \em N/A} & 35 \\
		\hline
		% apodolsk
		Group 5 & 50 & 15 & {\footnotesize \em N/A} & 15 & {\footnotesize \em N/A} & {\footnotesize \em N/A} & {\footnotesize \em N/A} & {\footnotesize \em N/A} \\
		\hline
		Former TA & 40 & 20 & 3 & 95 & {\footnotesize \em N/A} & 30 & 158 & 30 \\
		\hline
		\hline
		{\bf Average } & {\bf 22 } & {\bf 26.33 } & {\bf 9.43 } & {\bf 48.5 } & {\bf 31.25 } & {\bf 26 } & {\bf 119 } & {\bf 36.43} \\
		\hline
	\end{tabular}
	\end{center}
	\caption{Self-reported times spent by users on each phase while using Landslide. ``N/A'' indicates a student reported spending no time on a particular phase. ``Required'' indicates the sum of the times from the first four phases; ``refinement'' inducates the sum from the last two phases.
	``N/A'' indicates that a user did not report spending any time on a phase; the ``required'' field is ``N/A'' if any of its represented phases are, and ``N/A'' fields are not reflected in the averages.}
	\label{fig:student-times}
\end{table}
\begin{figure}[h]
	\center
	\includegraphics[width=0.5\textwidth]{graphs/studence-pie.pdf}
	\caption{Average time breakdown for each phase of instrumenting a kernel with Landslide, based on student self-reported times.}
	\label{fig:student-pie}
\end{figure}

@@

Table~\ref{fig:student-times} presents the time breakdown among phases that students experienced while using Landslide, visualised in Figure~\ref{fig:student-pie}. On average, students spent two hours getting Landslide to work with their kernels, and slightly more than half an hour customising the search and/or studying found bugs on top of that.

@@

Table~\ref{fig:numbers} shows the time it takes to find each of these bugs using Landslide, configured with several different sets of decision points.
\revision{Figure~\ref{fig:number-graph} compares the total exploration time among each different decision set for the trials that did find bugs.}

\newcommand\bugnum[2]{\textcolor{BrickRed}{{\bf #1} {\scriptsize \em (#2)}}}
\newcommand\nobugnum[2]{\textcolor{Blue}{[{\em #1 {\scriptsize (#2)}}]}}

\begin{table}[t!]
	\begin{center}
	\footnotesize
	\begin{tabular}{|l||c|c|c|c|}
		\hline
		\multirow{2}{*}{\bf Kernel and test case} & \multicolumn{4}{c|}{\bf Decision points used} \\ % & \multirow{2}{*}{\bf Stress test} \\
		\cline{2-5}
		& {\bf default} & {\bf lock} & {\bf unlock} & {\bf both} \\
		\hline\hline
		POBBLES vanish/vanish (a) & \nobugnum{31.8}{0.6} & \bugnum{57.1}{1.7} & $\infty$ & $\infty$ \\
		\hline
		POBBLES vanish/vanish (b) & \nobugnum{32.0}{0.4} & \bugnum{51.5}{2.1} & \nobugnum{8057.9}{336.9} & $\infty$ \\
		\hline
		POBBLES wait/wait & \bugnum{23.3}{0.7} & \bugnum{27.9}{0.8} & \bugnum{27.9}{2.1} & \bugnum{41.6}{1.4} \\
		\hline
		POBBLES thread\_fork/vanish & \bugnum{22.0}{0.6} & \bugnum{37.4}{1.1} & \bugnum{27.6}{0.5} & \bugnum{72.0}{2.6} \\
		\hline
		LudicrOS vanish/vanish & \nobugnum{13.2}{0.2} & \bugnum{13.7}{0.7} & \bugnum{34.6}{1.1} & \bugnum{17.1}{0.3} \\
		\hline
		LudicrOS yield/vanish & \nobugnum{12.3}{0.3} & \bugnum{11.4}{0.4} & \nobugnum{27.4}{0.8} & \bugnum{11.7}{0.4} \\
		\hline
		\multicolumn{5}{c}{Key: \bugnum{seconds}{stddev} indicates bug found; \nobugnum{seconds}{stddev} indicates whole tree explored with no bug.} \\
	\end{tabular}
	\end{center}
	\caption{Comparison of time taken (in seconds) to find bugs using Landslide with various decision sets:
	the default set, consisting only of voluntary reschedules (Section~\ref{sec:components-arbiter}); and using custom decision points in addition to the default set: calls to \texttt{mutex\_lock}, calls to \texttt{mutex\_unlock}, and both.
	All numbers represent the average from 5 trials, with the standard deviations given in parentheses. ``$\infty$'' indicates that Landslide's search did not finish (after 8 hours).
	}
	\label{fig:numbers}
\end{table}
\revision{
\begin{figure}[h]
	\center
	\includegraphics[width=0.7\textwidth]{graphs/decisions.pdf}
	\caption{Bar graph visualisation of the exploration time in trials that did find bugs. The decision trees that did not expose bugs, and the ones that timed out, are not shown.}
	\label{fig:number-graph}
\end{figure}
}

The experimental set-up is as follows:

@@

%The stress test experimental set-up is as follows:
%
%\begin{itemize}
%	\item A wrapper program runs 1024 simultaneous copies of the same test program that Landslide uses. Whenever one of them exits, the wrapper starts a new one.
%	\item All stress test trials are on on the ``412 lab crashbox'' (3.2 GHz Pentium D).
%	\item The kernel and test case are configured to not print any messages during the test
%\end{itemize}

@@

\newcommand\bugtree[1]{\textcolor{BrickRed}{\bf #1}}
\newcommand\nobugtree[1]{\textcolor{Blue}{[{\em #1}]}}
\begin{table}[t!]
	\begin{center}
	\footnotesize
	\begin{tabular}{|l|l||c|c|c|c|}
		\hline
		\multirow{2}{*}{\bf Kernel and test case} & \multirow{2}{*}{\bf Property of tree} & \multicolumn{4}{|c|}{\bf Decision points used} \\
		\cline{3-6}
		& & \bf default & \bf lock & \bf unlock & \bf both \\
		\hline\hline
		\multirow{4}{*}{POBBLES vanish/vanish (a)} & Decision points & \nobugtree{56} & \bugtree{1296} & N/A & N/A \\
		& Total backtracks   & \nobugtree{16} & \bugtree{376} & N/A & N/A \\
		& Average branch depth & \nobugtree{5} & \bugtree{19} & N/A & N/A \\
		\hline
		\multirow{4}{*}{POBBLES vanish/vanish (b)} & Decision points & \nobugtree{56} & \bugtree{1295} & \nobugtree{382071} & N/A \\
		& Total backtracks   & \nobugtree{16} & \bugtree{376} & \nobugtree{112706} & N/A \\
		& Average branch depth & \nobugtree{5} & \bugtree{17} & \nobugtree{16} & N/A \\
		\hline
		\multirow{4}{*}{POBBLES wait/wait} & Decision points & \bugtree{23} & \bugtree{102} & \bugtree{74} & \bugtree{378} \\
		& Total backtracks   & \bugtree{4} & \bugtree{17} & \bugtree{12} & \bugtree{56} \\
		& Average branch depth & \bugtree{6} & \bugtree{10} & \bugtree{9} & \bugtree{14} \\
		\hline
		\multirow{4}{*}{POBBLES thread\_fork/vanish} & Decision points & \bugtree{24} & \bugtree{394} & \bugtree{273} & \bugtree{2269} \\
		& Total backtracks   & \bugtree{5} & \bugtree{70} & \bugtree{56} & \bugtree{410} \\
		& Average branch depth & \bugtree{5} & \bugtree{16} & \bugtree{14} & \bugtree{23} \\
		\hline
		\multirow{4}{*}{LudicrOS vanish/vanish} & Decision points & \nobugtree{10} & \bugtree{16} & \bugtree{141} & \bugtree{42} \\
		& Total backtracks   & \nobugtree{2} & \bugtree{3} & \bugtree{48} & \bugtree{10} \\
		& Average branch depth & \nobugtree{2} & \bugtree{7} & \bugtree{9} & \bugtree{14} \\
		\hline
		\multirow{4}{*}{LudicrOS yield/vanish} & Decision points & \nobugtree{8} & \bugtree{5} & \nobugtree{149} & \bugtree{7} \\
		& Total backtracks   & \nobugtree{1} & \bugtree{0} & \nobugtree{43} & \bugtree{0} \\
		& Average branch depth & \nobugtree{2} & \bugtree{0} & \nobugtree{9} & \bugtree{0} \\
		\hline
	\end{tabular}
	\end{center}
	\caption{Information about the decision trees explored when finding bugs. As in the previous table, each test case was run with the four different sets of decision points. ``\nobugtree{X}'' means the tree was completely explored because Landslide did not find a bug in that configuration. ``\bugtree{X}'' reflects the portion of the tree that was explored before a bug was found.}
	\label{fig:trees}
\end{table}
